---
title: "whitebarn fall baseline samples 2023"
format: pdf
editor: visual
editor_options: 
  chunk_output_type: console
---

```{r echo=FALSE,include=FALSE}
source("whitebarn_nir_summary.R")

```

# NIRS

forage quality estimated using near infrared spectroscopy (NIRS) Perten DA7250 with 'hay' equation in small white static plastic cup.

machine struggled to confidently predict neutral detergent fiber digestibility so I am not reporting those values nor relative forage quality.

protein, adf (acid detergent fiber) and ndf (neutral detergent fiber) expressed as percent of dry matter.

rfv.sdsu is relative feed value (RFV) and has no units.

# Straw bale

Over a dozen cores of straw bales were sampled at white barn during fall sampling. We scanned 8 sub-samples to predict forage quality.

Protein averaged 5.9 %, RFV averaged 63.

```{r echo=F}
# just data from straw bale sample
dat8 %>% 
  select(protein,adf,ndf,rfv.sdsu) %>% 
  summary()

```

# Plot data

We collected fall baseline samples from 3 locations in the field. From each location, 3 samples within the exclosure and 3 samples outside. Therefore we had 3\*(3+3) total samples from the field, each sample was scanned 3 times after grinding, resulting in 54 total data points for forage quality.

```{r echo=FALSE}
dat9 %>% 
  select(-c(sample_type,sample_location,plot,Experiment,exclosure)) %>% 
  summary()
```

```{r echo=F, message=FALSE}
# data without straw bale
dat9 %>% 
  ggplot(aes(protein)) +
  stat_bin()

dat9 %>% 
  ggplot(aes(rfv.sdsu)) +
  stat_bin()
```

# Comparisons

```{r echo=FALSE}
dat9 %>% 
  # glimpse()
  ggplot(aes(sample_location,protein,
             fill=exclosure)) +
  geom_boxplot()

dat9 %>% 
  # glimpse()
  ggplot(aes(sample_location,rfv.sdsu,
             fill=exclosure)) +
  geom_boxplot()
```



# Conclusions

These are fall baseline samples of intermediate wheatgrass. 
We expect good quality forage. This would be protein around 15% and relative feed value of 150.
We expect normally distributed data
We do not expect any differences by exclosure (inside vs outside). 
