---
title: "RALLF Forage Quality"
subtitle: "preliminary report"
date: "2023 February"
author: "Jesse Puka-Beals"
output:
  pdf_document:
  # word_document
    toc: true
editor_options: 
  chunk_output_type: console
format: 
  pdf: 
    colorlinks: false
    include-in-header: 
      text: |
        \usepackage{hyperref}
---

### To do

Put in dates of harvests. 

Recreate the RFQ equation to understand what will happen to RFQ prediction as NDFD is adjusted

Update this report after we finish scanning 2022 data. We found more data on 7March for 2022-3rd cut. 

Validate the predicted forage quality parameters with wet chemsity. If machine is underestimating ADF and NDF, this will explain the high RFV values. The NDFD values need to also be validated. When we predicted NDFD for IWG in FIG the values were reasonable (50%), but for alfalfa they are unreasonable (16%). We need to consult with a lab familiar with NIRS for both conventional and reduced lignin alfalfa. 

Look at predicted lignin from NIRS?

\newpage
### Expected results

Alfalfa varieties (HarvXtra) will have lower lignin content and higher NDFD than conventional varieties. The RFV should be higher for HarvXtra varieties. 

Alfalfa with lower intervals between cuttings (35-day vs 45-day ) will have higher forage quality (RFV, CP).


### Discussion topics

*RFV vs. RFQ*

If we are only making comparisons among alfalfa cuttings, let's use RFV not RFQ. RFQ relies on NDFD, which is much harder to predict with NIRS. There is lot of variability in NDFD in wet chemsity due to variability in rumen fluid composition from donor cows, this variability extends into the NIRS database for NDFD, which reduces the accuracy of the NIRS predictions. RFQ is useful when comparing between mixtures of legumes and grasses, but since we are only making comparisons between alfalfa cuttings, RFV allows us to rely on just ADF and NDF, which NIRS performs better at predicting. 

*NIRS predictions*

Do we trust the predicted values of the NIRS for this experiment?

Is the hay equation sufficient for making comparisons between HarvXtra vs. conventional alfalfa forage quality?

Why is the RFV prediction so high for alfalfa? Did the hail storm in May2022 impact forage quality? Perhaps people put more leaves in the forage quality subsample then stems, causing much higher RFV. 

Why is neutral detergent digestibility so low? We need to be skeptical of RFQ so long as we are skeptical of NDFD, even though RFQ is correctly correlated with RFV. 

Why is variance so low in the NIRS prediction? Is there an issue with how data is being read in or processed causing pseudoreplication?

### Data availability

\href{https://github.com/jpukabeals/nir/tree/main/rallf}{Click Me for Github Repository}
*https://github.com/jpukabeals/nir/tree/main/rallf*

```{r, echo=F, include=F}
knitr::opts_chunk$set(
  echo = F,
  warning = F, 
  message = F,
  comment = NA
)

library(tidyverse)
library(tidyr)

rm(list = ls())

theme_set(theme_bw())

```

```{r, include=F}
setwd("~/R projects/nir/rallf")

source("rallf_keys.R")

source("rallf_nir_22Feb2023.R")

```

\newpage

# Data we have

```{r}

dat5 %>% 
  dplyr::select(code,year,plot,rfq.legume) %>% 
  left_join(key3) %>%
  # glimpse()
  full_join(key2) %>% 
  # glimpse()
  left_join(key1) -> dat2


dat2 %>% 
  group_by(
    year,site,intensity,cut, harvest_code
  ) %>% 
  tally() %>% 
  filter(year<2023) %>% 
  mutate(n = na_if(n,1),
         n = replace_na(n,0)) %>% 
  arrange(harvest_code) %>% 
  filter(harvest_code<15) %>% 
  mutate(newcode = paste(year,harvest_code,sep="_")) %>% 
  filter(newcode != "2021_11" &
           newcode != "2021_12" &
           newcode != "2021_13" &
           newcode != "2021_14") %>% 
  dplyr::select(-newcode) %>% 
  mutate(n_expected = 24,
         n_missing = n_expected-n) %>% 
  dplyr::select(-c(n_expected,harvest_code)) %>% 
  knitr::kable(
    caption = "NIRS data"
  )

dat2 %>% 
  group_by(
    year,site,intensity,cut, harvest_code
  ) %>% 
  tally() %>% 
  filter(year<2023) %>% 
  mutate(n = na_if(n,1),
         n = replace_na(n,0)) %>% 
  arrange(harvest_code) %>% 
  filter(harvest_code<15) %>% 
  mutate(newcode = paste(year,harvest_code,sep="_")) %>% 
  filter(newcode != "2021_11" &
           newcode != "2021_12" &
           newcode != "2021_13" &
           newcode != "2021_14") %>% 
  dplyr::select(-newcode) %>% 
  mutate(n_expected = 24,
         n_missing = n_expected-n) %>% 
  dplyr::select(-c(n_expected,harvest_code)) %>% 
  filter(n_missing!=0) %>%
  knitr::kable(
    caption = "Missing NIRS data"
  )

```

As seen in tables, we have missing $24+24+1+14+22 = 85$ data points. The two sets of 24 data points are for the first cut at rosemount in 2022, which have not been scanned as of 15Feb2023. The remaining missing data may have been tossed due to mold.

^^ we found the 3rd cut data 

\newpage

```{r}

dat5 %>% 
  # colnames()
  dplyr::select(code,year,plot,
                CP,ADF,NDF,NDFD,rfv,
                rfq.legume) %>% 
  left_join(key3) %>% 
  left_join(key2) %>% 
  left_join(key1) -> dat1

```

# Predicting forage quality parameters with NIRS

We are predicting alfalfa forage quality parameters using near infrared scanning with a generalist equation designed for all hay. Some labs use an alfalfa equation and some labs have different alfalfa equations for conventional vs. reduced lignin alfalfa. All predictions should be validated with wet chemistry, though NIRS predictions without recent validation can still be useful in estimating relative differences between treatments.

Soil contamination can also cause issues with forage quality predictions.

Expected forage quality parameters for alfalfa were easy to obtain for CP ADF NDF and RFV, they were not easy to obtain for RFQ. In general, I found that RFQ can range from 50 to 250, with most data between 100 and 200 and an average of 150.

```{r Dairyland predicted values}
c(
  "CP",
  "ADF",
  "NDF",
  "Lignin"
) -> prmtr

c(
  20,
  31,
  38,
  6
) -> mdn

c(
  14,
  20,
  35,
  4
) -> mn

c(
  26,
  43,
  52,
  8
) -> mx

tibble(
  prmtr,
  mdn,
  mn,
  mx
) %>% 
  mutate(
    source="dairyland labs",
    .before=prmtr
  ) %>% 
  mutate(range = paste(mn,mx,sep = "-")) %>% 
  rename(
    median = mdn,
    parameter = prmtr
  ) %>% 
  dplyr::select(-c(mn,mx,source)) %>% 
  rename_all(str_to_title) %>% 
  knitr::kable(
    caption = "Expected forage quality ranges from Dairyland Labs"
  )

```

```{r predicted ranges aggregated}

# https://docs.google.com/spreadsheets/d/1AqVcEwv3PlHVcR5Vv5UqUliUcqOZj6dNMvK-FAArONg/edit#gid=1646351822

read.csv(
  "alfalfa_forageQuality_expectedValues.csv"
) -> dat_qual


dat_qual %>% 
  rename(
    "Grade" = quality.category,
  ) %>% 
  rename_all(str_to_title) %>% 
  rename_with(
    str_to_upper,
    4:8
  ) %>% 
  knitr::kable(
    caption= "Expected forage quality ranges from UMN, SDSU, Univeristy of Georgia and Dairyland"
  )

```

```{r}

dat_qual %>% 
  dplyr::filter(stage=="all") %>% 
  dplyr::select(-c(maturity,stage,quality.category)) %>% 
  rename_all(toupper) %>% 
  mutate(type = "expected") -> d1


dat1 %>% 
  summarise(
    across(
      c(CP,ADF,NDF,rfv,rfq.legume),
      mean
    )
  ) %>% 
  rename(
    RFQ = rfq.legume,
    RFV = rfv
  ) %>% 
  mutate_all(
    ~round(.x,0)
  ) %>% 
  mutate(
    type = "observed"
  ) %>% 
  mutate_all(as.character) %>% 
  bind_rows(d1) %>% 
  relocate(type,.before=CP) -> d2 

# d2 %>% 
#   knitr::kable()

c(
  20,
  30,
  40,
  150,
  150
) -> expctd

as_tibble(cbind(nms = names(d2), t(d2))) %>% 
  rename(
    Observed = V2,
    `Expected Range` = V3,
    `Forage Quality Parameter` = nms
  ) %>% 
  slice(-1) %>% 
  bind_cols(expctd) %>% 
  rename(Expected = "...4") %>% 
  relocate(Expected,.after=Observed) %>% 
  knitr::kable(
    caption = "Observed vs. expected forage quality in RALLF"
  )

```

\newpage

# Summary tables

```{r FQ by rep}

dat1 %>% 
  group_by(rep) %>% 
  rename_with(
    .cols = c(1:3, 11:15),
    str_to_title
  ) %>% 
  rename(
    RFV = rfv,
    RFQ = rfq.legume
  ) %>% 
  summarise(
    across(c(CP,ADF,NDF,NDFD,RFV,RFQ),
           .fns=mean)
  ) %>% 
  mutate(
    across(c(CP,ADF,NDF,NDFD,RFV,RFQ),
           .fns=~round(.x,0))
  ) %>% 
  knitr::kable(
    caption = "Forage quality by rep"
  )

```

```{r FQ by site}
dat1 %>% 
  group_by(site) %>% 
  rename_with(
    .cols = c(1:3, 11:15),
    str_to_title
  ) %>% 
  rename(
    RFV = rfv,
    RFQ = rfq.legume
  ) %>% 
  summarise(
    across(c(CP,ADF,NDF,NDFD,RFV,RFQ),
           .fns=mean)
  ) %>% 
  mutate(
    across(c(CP,ADF,NDF,NDFD,RFV,RFQ),
           .fns=~round(.x,0))
  ) %>% 
  knitr::kable(
    caption = "Forage quality by site"
  )

```

```{r FQ by year}
dat1 %>% 
  group_by(year) %>% 
  rename_with(
    .cols = c(1:3, 11:15),
    str_to_title
  ) %>% 
  rename(
    RFV = rfv,
    RFQ = rfq.legume
  ) %>% 
  # colnames()
  summarise(
    across(c(CP,ADF,NDF,NDFD,RFV,RFQ),
           .fns=mean)
  ) %>% 
  mutate(
    across(c(CP,ADF,NDF,NDFD,RFV,RFQ),
           .fns=~round(.x,0))
  ) %>% 
  knitr::kable(
    caption = "Forage quality by year"
  )
```

```{r FQ by cut}
dat1 %>% 
  group_by(cut) %>% 
  rename_with(
    .cols = c(1:3, 11:15),
    str_to_title
  ) %>% 
  rename(
    RFV = rfv,
    RFQ = rfq.legume
  ) %>% 
  summarise(
    across(c(CP,ADF,NDF,NDFD,RFV,RFQ),
           .fns=mean)
  ) %>% 
  mutate(
    across(c(CP,ADF,NDF,NDFD,RFV,RFQ),
           .fns=~round(.x,0))
  ) %>% 
  knitr::kable(
    caption = "Forage quality by cutting"
  )
```

```{r FQ by cutting intensity}
dat1 %>% 
  group_by(intensity) %>% 
  rename_with(
    .cols = c(1:3, 11:15),
    str_to_title
  ) %>% 
  rename(
    RFV = rfv,
    RFQ = rfq.legume
  ) %>% 
  summarise(
    across(c(CP,ADF,NDF,NDFD,RFV,RFQ),
           .fns=mean)
  ) %>% 
  mutate(
    across(c(CP,ADF,NDF,NDFD,RFV,RFQ),
           .fns=~round(.x,0))
  ) %>% 
  knitr::kable(
    caption = "Forage quality by harvest schedule"
  )

```

\newpage
```{r FQ by variety}

dat1 %>% 
  group_by(variety) %>% 
  rename_with(
    .cols = c(1:3, 11:15),
    str_to_title
  ) %>% 
  rename(
    RFV = rfv,
    RFQ = rfq.legume
  ) %>% 
  summarise(
    across(c(CP,ADF,NDF,NDFD,RFV,RFQ),
           .fns=mean)
  ) %>% 
  mutate(
    across(c(CP,ADF,NDF,NDFD,RFV,RFQ),
           .fns=~round(.x,0))
  ) %>% 
  knitr::kable(
    caption = "Forage quality by variety"
  )

```

\newpage

# Histograms

```{r RFQ}

dat1 %>% 
  # colnames()
  ggplot(aes(rfq.legume)) +
  stat_bin() +
  labs(
    x="RFQ"
  )

```

We expect an RFQ of around 150, we observed an RFQ of `r round(mean(dat1$rfq.legume),0)`

```{r}
dat1 %>% 
  # colnames()
  ggplot(aes(rfv)) +
  stat_bin() +
  labs(
    x="RFV"
  )
```

We expect an RFV of around 150, we observed an RFV of `r round(mean(dat1$rfv),0)`. Some samples were predicted to have an RFV greater than 300!

```{r}
dat1 %>% 
  # colnames()
  ggplot(aes(rfv,
             fill = factor(cut))) +
  stat_bin() +
  labs(
    x="RFV"
  ) +
  geom_vline(xintercept = 200) +
  facet_wrap(~site*year)
```

With the exception of Rosemount in 2022, all site-years had multiple cuts with predicted RFV over 200. First cuts tended to have the highest quality, which is expected.

```{r}
dat1 %>% 
  ggplot(aes(CP)) +
  stat_bin() 
```

We expect a crude protein of around 20%, we observed a CP of `r round(mean(dat1$CP),0)`

```{r}
dat1 %>% 
  ggplot(aes(ADF)) +
  stat_bin()

```

We expect an ADF of around 30%, we observed an ADF of `r round(mean(dat1$ADF),0)`

```{r}
dat1 %>% 
  ggplot(aes(NDF)) +
  stat_bin()

```

We expect an NDF of around 40%, we observed an NDF of `r round(mean(dat1$NDF),0)`


```{r}
dat1 %>% 
  ggplot(aes(NDFD)) +
  stat_bin()

```

For neutral detergent digestibility over 48 hours, we expect values near 50%, we predicted values of `r round(mean(dat1$NDFD),0)`

\newpage

# Analysis

## Relative forage quality

Since relative feed value and relative forage quality are strongly correlated, the results of the anlaysis are the same. 

```{r echo=T}

dat1 %>% 
  lm(
    rfq.legume~site*year*variety*intensity*cut,.
  ) %>% 
  anova() %>%
  filter(`Pr(>F)` < 0.05)
# site:year:intensity
# year:intensity:cut

# no effect from variety
# some effect from intensity

```

\newpage

```{r}
dat1 %>% 
  ggplot(aes(cut,rfq.legume,
             col=intensity)) +
  stat_summary(  ) +
  facet_wrap(~site*year) +
  labs(y="RFQ")
```

Similar pattern across years and sites; RFQ is similar for the first cut and then the second cut the RFQ is lower for the 45-day cuts than the 35-day cuts. 

Why is variance so low? I can barely see error bars


```{r checking variance, echo=T}
# is their psuedoreplication/duplication happening causing variance to be low?
# this could be a mistake I made during a left_join()
dat1 %>% 
  filter(
    site=="rosemount" &
      year == "2021" &
      intensity == "45-day" &
      cut=="2"
  ) %>% 
  distinct(rfq.legume) %>% 
  .$rfq.legume 
# note that there are no duplicated values of RFQ

```

Let's look at standard deviation instead of standard error

```{r}
dat1 %>% 
  ggplot(aes(cut,rfq.legume,
             col=intensity)) +
  stat_summary(
    fun.data = mean_sdl
  ) +
  facet_wrap(~site*year) +
  labs(y="RFQ + Standard deviation")
```

There is simply a large difference between standard error and standard deviation, likely because n() is larger than usual (n=24) so standard error really reduces the visual representation of variance. 

\newpage

## Crude protein

```{r echo=T}

dat1 %>% 
  lm(
    CP~site*year*variety*intensity*cut,.
  ) %>% 
  anova() %>%
  filter(`Pr(>F)` < 0.05)
# 4-way interaction

# no effect from variety
# some effect from intensity

```

\newpage

```{r}
dat1 %>% 
  ggplot(aes(cut,CP,
             col=intensity)) +
  stat_summary() +
  facet_wrap(~site*year) +
  labs(y="CP")
```

Similar pattern across years and sites, same as RFQ; CP is similar for the first cut and then the second cut the CP is lower for the 45-day cuts than the 35-day cuts. 

Same as with RFQ, summarizing with standard error shows much less variation around means than standard deviation. 

\newpage
# Diving deeper

## Relative feed value and ADF

RFV is the only parameter where we are outside of our expected range. We expect around 150 and we are averaging 200. We have some RFV values over 300. 

This may be due to an issue with the equation

Let's look at the RFV equation I used

```{r}
print(calc.rfq.rfv.legume)
```

So I used ...

Digestible dry matter\
$DDM.rfv = 88.9-(0.779*ADF)$

Dry matter intake\
$DMI.rfv = 120/NDF$

Relative Forage Value\
$RFV = (88.9-(0.779*ADF))*(120/NDF)/1.29$

These equations are correct. 


```{r testing RFV equation, include=F}

# low range of fiber
ADF <- 27
NDF <- 34
# RFV was 185

ADF <- 29
NDF <- 36
# RFV was 171

# expected values
ADF <- 30
NDF <- 40
# RFV=150

# observed values
ADF <- 24
NDF <- 33
# RFV = 200

# ratio of ADF:NDF weird?
33/24
# 1:1.3
40/30
# 1:1.3

# the ratio of ADF to NDF is what we expect
# for 1ADF there's 1.3 NDF

(88.9-(0.779*ADF))*(120/NDF)/1.29
```

Let's look at the samples with the highest RFV values

```{r}
dat1 %>% 
  arrange(desc(rfv)) %>% 
  slice(1:5) %>% 
  dplyr::select(
    site,year,intensity,cut,variety,rep,
    CP,ADF,NDF,NDFD,rfv,rfq.legume
  ) %>% 
  # print()
  mutate(
    across(
      7:12,
      ~round(.x,0)
    )
  ) %>% 
  knitr::kable(
    caption = "The samples with the highest estiamted relative feed value"
  )
  
```

The forage samples with the highest RFV are from st paul in 2022. The ADF is 16% and the NDF is 20%. A forage sample with these values would have a relative feed value of 355, so the equation is not incorrect. The question is whether alfalfa can be this high quality. 

Can an alfalfa sample be 16% ADF and 20% NDF? 

It's notable that the crude protein is also very high, 29%, where we would normally expect 20%. This supports the idea that this was young vegetative leafy alfalfa. 

It seems very possible that our predicted ADF and NDF values simply need a bias adjustment. If this is the case, they can still be used to determine relative differences between treatments, but will need a wet chemistry calibration before they can be used to determine the actual ADF or NDF values. 

\newpage
## NDFD 

Neutral detergent fiber digestibility estimates the amount of digestion that will occur within the rumen and this varies by ruminant. The most common durations are 24 hours for cattle, 30 hours for sheep goats and horses, 48 hours for swine. There are also very long digestions (72, 120 or 240), which are meant to determine how much fiber is indigestible. 

We predict neutral detergent fiber digestibility over 48 hours. This is still an appropaite duration, but it can be harder to detect differences in forage digestibility with longer durations. We expect an NDFD range of 30-50% for alfalfa based on data from Dairyland and Forage Genetics International. 

```{r expected lignin and NDFD values}

c(
  "HarvXtra",
  "Conventional"
) ->alf_var

c(
  "35-day",
  "45-day"
) -> intensity


c(
  5,4.5,6.5,5.7
) -> lgnn

c(
  40,49,35,43
) -> ndfd_p

expand_grid(
  alf_var,
  rev(intensity)
) %>% 
  bind_cols(
    lgnn,ndfd_p
  ) %>% 
  rename(
    Variety = alf_var,
    Intensity = `rev(intensity)`,
    `Lignin (%)` = ...3,
    `NDFD (%)` = ...4
  ) %>% 
  knitr::kable(
    caption = "Expected lignin and NDFD values of different varieties and harvest schedules"
  )
# https://www.foragegenetics.com/FGI/media/Images/PDF/FGI_Comparison_Between_HarvXtra_Alfalfa_and_Varieties_Selected_for_High-Quality_White_Paper_1.pdf
```

```{r, include=F}
dat1 %>% 
  # group_by(variety,intensity) %>% 
  # summarise(NDFD = mean(NDFD)) %>% 
  mutate(Variety = str_sub(variety,1,2)) %>% 
  group_by(Variety,intensity) %>% 
  summarise(NDFD_48hr = mean(NDFD),
            sd = sd(NDFD)) %>% 
  # glimpse()
  mutate(
    across(
      c(NDFD_48hr,sd),
      .fns = ~round(.x,1)
    )
  ) %>% 
  mutate(
    Variety = factor(Variety),
    Variety = fct_recode(Variety,
                         HarvXtra = "HX",
                         Conventional = "RR")
  ) %>%  
  knitr::kable(
    caption = "Predicted NDFD mean and standard deviation observed in RALLF between HarvXtra (HX) and conventional (RR) alfalfa"
  )
```

```{r}

dat1 %>% 
  # group_by(variety,intensity) %>% 
  # summarise(NDFD = mean(NDFD)) %>% 
  mutate(Variety = str_sub(variety,1,2)) %>% 
  group_by(Variety,intensity) %>% 
  summarise(NDFD_48hr = mean(NDFD),
            sd = sd(NDFD)) %>% 
  # glimpse()
  mutate(
    across(
      c(NDFD_48hr,sd),
      .fns = ~round(.x,1)
    )
  ) %>% 
  mutate(
    Variety = factor(Variety),
    Variety = fct_recode(Variety,
                         HarvXtra = "HX",
                         Conventional = "RR")
  ) %>% 
  dplyr::select(-sd) %>% 
  mutate(type = "NDFD_observed") %>% 
  rename(`NDFD (%)` = NDFD_48hr,
         Intensity = intensity) -> dm1


expand_grid(
  alf_var,
  rev(intensity)
) %>% 
  bind_cols(
    lgnn,ndfd_p
  ) %>% 
  rename(
    Variety = alf_var,
    Intensity = `rev(intensity)`,
    `Lignin (%)` = ...3,
    `NDFD (%)` = ...4
  ) %>% 
  mutate(type = "NDFD_expected") %>% 
  dplyr::select(-`Lignin (%)`) %>% 
  bind_rows(dm1) %>% 
  relocate(type,.before=Variety) %>% 
  # print()
  pivot_wider(
    names_from=type,
    values_from = `NDFD (%)`
  ) %>% 
  knitr::kable(
    caption = "Predicted NDFD values did not match with expected values"
  )

```

We observed around 15% NDFD, which very low. 

We expected as time between cuttings decreased (35-day harvest schedule), NDFD would increase, however we observed a decrease in NDFD for both the HarvXtra and Conventional varieties

We expected greater NDFD in HarvXtra vs conventional, but observed lower NDFD. 

One possible explanation is that our NDFD prediction needs a new bias or perhaps a new equation for alfalfa NDFD as it differs from grass NDFD. 

Another explanation is that we need a different equation for HarvXtra alfalfa vs. conventional alfalfa. 

The predicted NIRS values for NDFD should be viewed with skepticism.


```{r}

dat1 %>% 
  # glimpse()
  summarise(
    max(NDFD),
    min(NDFD),
    median(NDFD)
  ) %>% 
  knitr::kable(
    caption = "Neutral detergent digestibility ranged from 4 to 31%. This is very low and should be validated with wet chemistry"
  )
```

Does the low NDFD values make the RFQ values lower, causing the RFQ and RFV to not be well correlated?

```{r}
library(corrplot)

# dat1 %>% 
#   dplyr::select(rfv,rfq.legume) %>% 
#   cor() %>% 
#   corrplot()
# ok, RFQ and RFV are very positively correlated

dat1 %>% 
  dplyr::select(rfv,rfq.legume, ADF, NDF,CP,NDFD) %>% 
  cor() %>% 
  corrplot()
# As NDF and ADF increase, so does NDFD, and RFV and RFQ goes down. 
  
```

Yes, RFV and RFQ are strongly positively correlated as they should be. 

Other forage quality parameters are correlated as expected with the exception of NDFD. 

As forages mature, CP should decrease and NDFD should decrease, therefore they should be positively correlated, but instead they are negatively correlated. Furthermore, as NDFD increases RFQ should increase, but they are also negatively correlated. 

\newpage
### RFQ

Let's see how RFQ values would be if NDFD was near 60 instead of 6.

If they are closer to 200, then this would mean we likely need to bias the NDF and ADF (they are too low) and bias the NDFD (too low as well). When we increase these values, the RFV and RFQ will go down to more normal levels


```{r, include=F}
print(calc.rfq.rfv.legume)

```

Tried and failed to recreate RFQ equation

```{r, include=F}
# rfq.legume=DMI.legume*TDN.legume/1.23,
# DMI.legume=(120/NDF) + (NDFD - 45) * .374/1350*100,
# TDN.legume=(NFC*.98)+(CP*.93)+(FA*.97*2.25)+(NDFn*NDFD/100)-7,
# NFC=100-((0.93*NDF)+CP+EE+Ash),
# NDFn=NDF*0.93, 


NDF <- 30
CP <- 20
EE <- 2.05
Ash <- mean(dat5$Ash)
FA <- 1.05

NDFD <- mean(dat5$NDFD)
NDFD <- 7


(120/NDF) + (NDFD - 45) * .374/1350*100*(100-((0.93*NDF)+CP+EE+Ash)*.98)+(CP*.93)+(FA*.97*2.25)+(NDF*0.93*NDFD/100)-7/1.23

# this equation seems to be wrong, because it predicts RFQ of 324, but it should be near 130. Not sure what is problem, but RFQ has a problem!

(120/NDF) + (NDFD - 45) * .374/1350*100*(100-((0.93*NDF)+CP+EE+Ash)*.98)+(CP*.93)+(FA*.97*2.25)+(NDF*0.93*NDFD/100)-7/1.23

```


```{r}

# dat5 %>% 
#   slice(1) %>% 
#   View()

CP=18
ADF = 20
NDF =30
NDFD =10
NDFD = 60
DM = 90
EE = 2
FA = 1
Ash = 10

100-((0.93*NDF)+CP+EE+Ash) -> NFC

NDF*.93 -> NDFn

# NDFDp
22.7+0.664*NDFD -> NDFDp

# TDN.legume
(NFC*.98) + (CP*.93) + (FA*.97*2.25) + (NDFn*NDFD/100) - 7 -> TDN.legume

# DMI.legume
(120/NDF) + (NDFD-45) * 0.374/1350*100 -> DMI.legume

# rfq.legume
DMI.legume * TDN.legume/1.23

# 

```


let's see what we are shooting for
```{r}

dat5$rfq.legume %>% 
  range() %>% 
  round(0)

```

\newpage
# A review of RFV and RFQ

In a similar way you want to know how much N is in a fertilizer, you want to know how much digestible food is in a forage. The old school way was relative feed value (RFV), which is simply a measure of fiber content. The higher the fiber (ADF and NDf), the lower the RFV, meaning there isn't as much digestible food in the forage. Among alfalfa cuttings that are both 1 ton of dry matter, buy the one with the higher RFv because it's more dense in digestible energy. The forage with a lower RFv has more undigestible lignin that's basically trash that you have to transport, feed, takes up space in the cows stomach during digestion and then becomes waste. So RFV is just a function of ADF and NDF and it works great at making comparisons between multiple cuttings of the same hay field. The problem with RFV is when you try to make a comparison between an alfalfa hay and an alfalfa-grass mixture. The problem is that grasses have more fiber (ADF and NDF), but their fiber is more digestible. So grasses get lower RFV values than they deserve.  

