---
title: "RALLF Forage Quality"
subtitle: "preliminary report"
date: "2023 February"
author: "Jesse Puka-Beals"
output:
  pdf_document
editor_options: 
  chunk_output_type: console
---

### To do

Update this report after we finish scanning 2022 data

Pull in lignin data from NIRS scanning?

### Discussion topics

Do we trust these predicted values?

Why is the RFV prediction so high for alfalfa? Is it a mistake with the equation being used? Is it because the ADF values are so low? And if so, why are the ADF values so low? Is is because of the hail storm and then when forage quality samples were taken people put more leaves in the bags then stems?

### Data availability

Need to make a new RALLF NIRS repository

Files and dependencies

rallf_keys.R
"rallf_key_plotTreatments.csv"
"rallf_key_harvestCode.csv"
"rallf_key_codeHarvest.csv"


rallf_nir_22Feb2023.R
"functions_nir.R"
"RALLF_nirData_20Feb2023.csv"

RALLF_nirReport.Rmd
"alfalfa_forageQuality_expectedValues.csv"



```{r, echo=F, include=F}
knitr::opts_chunk$set(
  echo = F,
  warning = F, 
  message = F,
  comment = NA
)

library(tidyverse)
library(tidyr)

rm(list = ls())

theme_set(theme_bw())

```

```{r, include=F}
source("rallf_keys.R")

source("rallf_nir_22Feb2023.R")

```


\newpage
# Data we have

```{r}

dat5 %>% 
  dplyr::select(code,year,plot,rfq.legume) %>% 
  left_join(key3) %>%
  # glimpse()
  full_join(key2) %>% 
  # glimpse()
  left_join(key1) -> dat2


dat2 %>% 
  group_by(
    year,site,intensity,cut, harvest_code
  ) %>% 
  tally() %>% 
  filter(year<2023) %>% 
  mutate(n = na_if(n,1),
         n = replace_na(n,0)) %>% 
  arrange(harvest_code) %>% 
  filter(harvest_code<15) %>% 
  mutate(newcode = paste(year,harvest_code,sep="_")) %>% 
  filter(newcode != "2021_11" &
           newcode != "2021_12" &
           newcode != "2021_13" &
           newcode != "2021_14") %>% 
  dplyr::select(-newcode) %>% 
  mutate(n_expected = 24,
         n_missing = n_expected-n) %>% 
  dplyr::select(-c(n_expected,harvest_code)) %>% 
  knitr::kable(
    caption = "NIRS data"
  )

dat2 %>% 
  group_by(
    year,site,intensity,cut, harvest_code
  ) %>% 
  tally() %>% 
  filter(year<2023) %>% 
  mutate(n = na_if(n,1),
         n = replace_na(n,0)) %>% 
  arrange(harvest_code) %>% 
  filter(harvest_code<15) %>% 
  mutate(newcode = paste(year,harvest_code,sep="_")) %>% 
  filter(newcode != "2021_11" &
           newcode != "2021_12" &
           newcode != "2021_13" &
           newcode != "2021_14") %>% 
  dplyr::select(-newcode) %>% 
  mutate(n_expected = 24,
         n_missing = n_expected-n) %>% 
  dplyr::select(-c(n_expected,harvest_code)) %>% 
  filter(n_missing!=0) %>%
  knitr::kable(
    caption = "Missing NIRS data"
  )

```

As seen in tables, we have missing 24+24+1+14+22 data points. 
The two sets of 24 data points are for the first cut at rosemount in 2022, which have not been scanned as of 15Feb2023. 
The remaining missing data may have been tossed due to mold. 

\newpage
# Preliminary analysis of data

```{r}

dat5 %>% 
  # colnames()
  dplyr::select(code,year,plot,
                CP,ADF,NDF,NDFD,rfv,
                rfq.legume) %>% 
  left_join(key3) %>% 
  left_join(key2) %>% 
  left_join(key1) -> dat1



```

## Predicted forage quality parameters by NIRS

We are predicting alfalfa forage quality parameters using near infrared red scanning with a generalist equation designed for all hay. Some labs use an alfalfa equation and some labs have different alfalfa equations for conventional vs. reduced lignin alfalfa. All predictions should be validated with wet chemistry, though NIRS predictions without recent validation can still be useful in  estimating relative differences between treatments. 

Soil contamination can also cause issues with forage quality predictions. 

Expected forage quality parameters for alfalfa were easy to obtain for CP ADF NDF and RFV, they were not easy to obtain for RFQ. In general, I found that RFQ can range from 50 to 250, with most data between 100 and 200 and an average of 150. 



```{r Dairyland predicted values}
c(
  "CP",
  "ADF",
  "NDF",
  "Lignin"
) -> prmtr

c(
  20,
  31,
  38,
  6
) -> mdn

c(
  14,
  20,
  35,
  4
) -> mn

c(
  26,
  43,
  52,
  8
) -> mx

tibble(
  prmtr,
  mdn,
  mn,
  mx
) %>% 
  mutate(
    source="dairyland labs",
    .before=prmtr
  ) %>% 
  mutate(range = paste(mn,mx,sep = "-")) %>% 
  rename(
    median = mdn,
    parameter = prmtr
  ) %>% 
  dplyr::select(-c(mn,mx,source)) %>% 
  rename_all(str_to_title) %>% 
  knitr::kable(
    caption = "Expected forage quality ranges from Dairyland Labs"
  )

```

```{r predicted ranges aggregated}

# https://docs.google.com/spreadsheets/d/1AqVcEwv3PlHVcR5Vv5UqUliUcqOZj6dNMvK-FAArONg/edit#gid=1646351822

read.csv(
  "alfalfa_forageQuality_expectedValues.csv"
) -> dat_qual


dat_qual %>% 
  rename(
    "Grade" = quality.category,
  ) %>% 
  rename_all(str_to_title) %>% 
  rename_with(
    str_to_upper,
    4:8
  ) %>% 
  knitr::kable(
    caption= "Expected forage quality ranges from UMN, SDSU, Univeristy of Georgia and Dairyland"
  )

```


```{r}

dat_qual %>% 
  dplyr::filter(stage=="all") %>% 
  dplyr::select(-c(maturity,stage,quality.category)) %>% 
  rename_all(toupper) %>% 
  mutate(type = "expected") -> d1


dat1 %>% 
  summarise(
    across(
      c(CP,ADF,NDF,rfv,rfq.legume),
      mean
    )
  ) %>% 
  rename(
    RFQ = rfq.legume,
    RFV = rfv
  ) %>% 
  mutate_all(
    ~round(.x,0)
  ) %>% 
  mutate(
    type = "observed"
  ) %>% 
  mutate_all(as.character) %>% 
  bind_rows(d1) %>% 
  relocate(type,.before=CP) -> d2 

# d2 %>% 
#   knitr::kable()

c(
  20,
  30,
  40,
  150,
  150
) -> expctd

as_tibble(cbind(nms = names(d2), t(d2))) %>% 
  rename(
    Observed = V2,
    `Expected Range` = V3,
    `Forage Quality Parameter` = nms
  ) %>% 
  slice(-1) %>% 
  bind_cols(expctd) %>% 
  rename(Expected = "...4") %>% 
  relocate(Expected,.after=Observed) %>% 
  knitr::kable(
    caption = "Observed vs. expected forage quality in RALLF"
  )

```



\newpage
# Tables

```{r FQ by rep}

dat1 %>% 
  group_by(rep) %>% 
  rename_with(
    .cols = c(1:3, 11:15),
    str_to_title
  ) %>% 
  rename(
    RFV = rfv,
    RFQ = rfq.legume
  ) %>% 
  summarise(
    across(c(CP,ADF,NDF,NDFD,RFV,RFQ),
           .fns=mean)
  ) %>% 
  mutate(
    across(c(CP,ADF,NDF,NDFD,RFV,RFQ),
           .fns=~round(.x,0))
  ) %>% 
  knitr::kable(
    caption = "Forage quality by rep"
  )

```

```{r FQ by site}
dat1 %>% 
  group_by(site) %>% 
  rename_with(
    .cols = c(1:3, 11:15),
    str_to_title
  ) %>% 
  rename(
    RFV = rfv,
    RFQ = rfq.legume
  ) %>% 
  summarise(
    across(c(CP,ADF,NDF,NDFD,RFV,RFQ),
           .fns=mean)
  ) %>% 
  mutate(
    across(c(CP,ADF,NDF,NDFD,RFV,RFQ),
           .fns=~round(.x,0))
  ) %>% 
  knitr::kable(
    caption = "Forage quality by site"
  )

```


```{r FQ by year}
dat1 %>% 
  group_by(year) %>% 
  rename_with(
    .cols = c(1:3, 11:15),
    str_to_title
  ) %>% 
  rename(
    RFV = rfv,
    RFQ = rfq.legume
  ) %>% 
  # colnames()
  summarise(
    across(c(CP,ADF,NDF,NDFD,RFV,RFQ),
           .fns=mean)
  ) %>% 
  mutate(
    across(c(CP,ADF,NDF,NDFD,RFV,RFQ),
           .fns=~round(.x,0))
  ) %>% 
  knitr::kable(
    caption = "Forage quality by year"
  )
```


```{r FQ by cut}
dat1 %>% 
  group_by(cut) %>% 
  rename_with(
    .cols = c(1:3, 11:15),
    str_to_title
  ) %>% 
  rename(
    RFV = rfv,
    RFQ = rfq.legume
  ) %>% 
  summarise(
    across(c(CP,ADF,NDF,NDFD,RFV,RFQ),
           .fns=mean)
  ) %>% 
  mutate(
    across(c(CP,ADF,NDF,NDFD,RFV,RFQ),
           .fns=~round(.x,0))
  ) %>% 
  knitr::kable(
    caption = "Forage quality by cutting"
  )
```

```{r FQ by cutting intensity}
dat1 %>% 
  group_by(intensity) %>% 
  rename_with(
    .cols = c(1:3, 11:15),
    str_to_title
  ) %>% 
  rename(
    RFV = rfv,
    RFQ = rfq.legume
  ) %>% 
  summarise(
    across(c(CP,ADF,NDF,NDFD,RFV,RFQ),
           .fns=mean)
  ) %>% 
  mutate(
    across(c(CP,ADF,NDF,NDFD,RFV,RFQ),
           .fns=~round(.x,0))
  ) %>% 
  knitr::kable(
    caption = "Forage quality by harvest schedule"
  )

```

```{r FQ by variety}

dat1 %>% 
  group_by(variety) %>% 
  rename_with(
    .cols = c(1:3, 11:15),
    str_to_title
  ) %>% 
  rename(
    RFV = rfv,
    RFQ = rfq.legume
  ) %>% 
  summarise(
    across(c(CP,ADF,NDF,NDFD,RFV,RFQ),
           .fns=mean)
  ) %>% 
  mutate(
    across(c(CP,ADF,NDF,NDFD,RFV,RFQ),
           .fns=~round(.x,0))
  ) %>% 
  knitr::kable(
    caption = "Forage quality by variety"
  )

```


\newpage
# Histograms

```{r RFQ}

dat1 %>% 
  # colnames()
  ggplot(aes(rfq.legume)) +
  stat_bin() +
  labs(
    x="RFQ"
  )

```

We expect an RFQ of around 150, we observed an RFQ of `r round(mean(dat1$rfq.legume),0)`

```{r}
dat1 %>% 
  # colnames()
  ggplot(aes(rfv)) +
  stat_bin() +
  labs(
    x="RFV"
  )
```

We expect an RFV of around 150, we observed an RFV of `r round(mean(dat1$rfv),0)`


```{r}
dat1 %>% 
  ggplot(aes(CP)) +
  stat_bin() 
```

We expect a crude protein of around 20%, we observed a CP of `r round(mean(dat1$CP),0)`

```{r}
dat1 %>% 
  ggplot(aes(ADF)) +
  stat_bin()

```

We expect an ADF of around 30%, we observed an ADF of `r round(mean(dat1$ADF),0)`

```{r}
dat1 %>% 
  ggplot(aes(NDF)) +
  stat_bin()

```

We expect an NDF of around 40%, we observed an NDF of `r round(mean(dat1$NDF),0)`

\newpage
# Analysis

```{r echo=T}

dat1 %>% 
  lm(
    rfq.legume~site*year*variety*intensity*cut,.
  ) %>% 
  anova() %>%
  filter(`Pr(>F)` < 0.05)
# site:year:intensity
# year:intensity:cut

# no effect from variety
# some effect from intensity

```


\newpage
```{r}
dat1 %>% 
  ggplot(aes(cut,rfq.legume,
             col=intensity)) +
  stat_summary() +
  facet_wrap(~site*year) +
  labs(y="RFQ")
```

Similar pattern across years and sites; RFQ is similar for the first cut and then the second cut the RFQ is lower for the 45-day cuts than the 35-day cuts. However, we are still missing some data and will collect more data. 